Linking assumptions and number of response options modulate inferred scalar implicature rate

### 2018-05-09 writeup.Rmd

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative priors **JD: Masoud is the uninformative prior part correct?**. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

The model provided evidence for the following effects: First, there was a main effect of trial type such that scalar trials resulted in greater implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between link and number of response options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a Weak link resulted in greater implicature rates. Finally, there was a three-way interaction between link, trial type, and number of response options (Mean Estimate = `r round(fixed_effects$Estimate[15],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[15],2)`, `r round(fixed_effects$X97.5.ile[15],2)`]). One interpretation of this interaction is that the difference between the weak and Strong link on scalar trials in the quinary task was smaller than on exhaustive trials, though we believe this is not too interesting, given that the binary reference level implicature estimate was lower for exhaustive trials in the first place. **JD: happy to throw out this last sentence** Crucially, both number of response options and link affect the inferred implicature rate.

>>>>

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative or weakly informative priors [^3]. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

[^3: To know more about the default priors of the "brms" package you can refer to [the brms package manual](ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf).]

The model provided evidence for the following effects: First, there was a main effect of trial type such that scalar trials resulted in greater implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between link and number of response options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak link resulted in greater implicature rates. Finally, there was a three-way interaction between link, trial type, and number of response options (Mean Estimate = `r round(fixed_effects$Estimate[15],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[15],2)`, `r round(fixed_effects$X97.5.ile[15],2)`]). One interpretation of this interaction is that the difference between the weak and strong link on scalar trials in the quinary task was smaller than on exhaustive trials, though we believe this is not too interesting, given that the binary reference level implicature estimate was lower for exhaustive trials in the first place. Crucially, both number of response options and link affect the inferred implicature rate.
